%\documentclass[11pt]{paper}
%\usepackage{fullpage}
%\usepackage{palatino}
%\usepackage{amsfonts,amsmath,amssymb}
%% \usepackage{graphicx}
%
%\usepackage{listings}
%\usepackage{textcomp}
%\usepackage{color}
%
%\definecolor{dkgreen}{rgb}{0,0.6,0}
%\definecolor{gray}{rgb}{0.5,0.5,0.5}
%\definecolor{mauve}{rgb}{0.58,0,0.82}
%
%\lstset{frame=tb,
%  language=R,
%  aboveskip=3mm,
%  belowskip=3mm,
%  showstringspaces=false,
%  columns=flexible,
%  basicstyle={\small\ttfamily},
%  numbers=none,
%  numberstyle=\tiny\color{gray},
%  keywordstyle=\color{blue},
%  commentstyle=\color{dkgreen},
%  stringstyle=\color{mauve},
%  breaklines=true,
%  breakatwhitespace=true,
%  tabsize=3
%}
%
%
%
%\ifx\pdftexversion\undefined
%    \usepackage[dvips]{graphicx}
%\else
%    \usepackage[pdftex]{graphicx}
%    \usepackage{epstopdf}
%    \epstopdfsetup{suffix=}
%\fi
%
%\usepackage{subfig}
%
%
%% This allows pdflatex to print the curly quotes in the
%% significance codes in the output of the GAM.
%\UseRawInputEncoding
%
%\begin{document}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Problem Set 7
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\pagestyle{empty}
%{\noindent\bf Spring 2023 \hfill Brandon~Parmanand}
%\vskip 16pt
%\centerline{\bf University of Central Florida}
%\centerline{\bf College of Business}
%\vskip 16pt
%\centerline{\bf QMB 6911}
%\centerline{\bf Capstone Project in Business Analytics}
%\vskip 10pt
%\centerline{\bf Solutions:  Problem Set \#7}
%\vskip 32pt
%\noindent
%% 
%\section{Data Description}
%
%This analysis follows the script \texttt{PS7.R} to produce a more accurate model for used tractor prices with the data from \texttt{homesales.dat} in the \texttt{Data} folder. 

I will revisit the recommended linear model.
I will further investigate nonlinear relationships
by incorporating another nonlinear but parametric specification
for variables.
This parametric analysis will be performed
using the Box-Tidwell framework
to investigate whether the value of these characteristics
are best described with parametric nonlinear forms. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Linear Regression Model}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Suggested Linear Regression Model}
A natural staring point is the recommended linear model
from previous which we can see in Table \ref{tab:reg_comp}. 

% 
\input{../Tables/reg_comp}
% 
There are houses that were built in 2021 so in order to account for this in the model, an edit was made to calculate age by taking 2022- Year Built. Additionally, Number of Beds and Numbers of Bathrooms were treated as factors in the model as they more so categorize the houses.
%
\clearpage


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \clearpage
\section{Nonlinear Specifications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\pagebreak
\subsection{The Box--Tidwell Transformation}

The Box--Tidwell function tests for non-linear relationships
to the mean of the dependent variable.
The nonlinearity is in the form of an
exponential transformation in the form of the Box-Cox
transformation, except that the transformation is taken
on the explanatory variables.


\subsubsection{Transformation of Lot Size}


Performing the transformation on the Lot Size variable
produces a modified form of the linear model. The exponentis significantly differnt from 0. WItrh a small positive value that suggest an increasing relationship then leveling off to a slower increase. 

\input{../Tables/bt_lot}



\subsubsection{Transformation of Age}


\input{../Tables/bt_age}

This coefficient is significantly different than 0. With a positive value that suggest not quite a linear relationship but it is an increasing relationship with it leveling off and increasing at a slower rate.  

\subsubsection{Transformation of Transit Score}


\input{../Tables/bt_trans}

The transit score coefficient is not statistically significant but it is close to 1 showing close to linear realationship.

Since a nonlinear relationship was detected with age and lot size,
I will next estimate a model
with nonlinearity in all three continuous variables.


\subsubsection{Transformation of All Three Continuous Variables}


\input{../Tables/bt_full}


The performance is similar to the other models with
forms of nonlinearity .



\pagebreak
\section{Linear Approximation of the Box--Tidwell Transformation}

I created three variables 
\texttt{bt\_trans\_log\_trans}, \texttt{bt\_age\_log\_age}, and \texttt{bt\_lot\_log\_lot}, 
all of which were created by a transformation of the form $f(x) = x\cdot\log(x)$. 
Table \ref{tab:reg_bt_lin} collects the results
of the set of models from the nonlinear approximation to the models with the three forms of nonlinearity.
Model 1 is the linear regression model with  
the approximation of the transformation applied to transit score. 
Models 2 and 3
have the same specification as the other one, 
except that the transit score variable is replaced with
the variables for age and lot size, respectively. 
The coefficient on \texttt{bt\_age\_log\_age}
is the most statistically significant. 
And the coefficient on \texttt{bt\_lot\_log\_lot} is also statistically signifcant. 
This implies, just as the Box-Tidwell statistic predicts, 
a nonlinear relationship exists for the value of age and lot size.
The transit score is not statistically significant as the other two
indicating that a linear relationship suffices for the decline in value from transit score.

\input{../Tables/reg_bt_lin}


\pagebreak
\section{Comparison of Candidate Models}

Comparing the three models, Age and Lot Size were both significantly signifcant with p less than $0.001$ but the model with the age transformation had the highest R squared. I created a variable \texttt{age\_bt}
by raising age to the optimal exponent 
$\hat{\lambda} = 0.49715$. 
Then, I included this variable in the place of 
the age variables a the linear regression model.
% 
Table \ref{tab:reg_sq_horse_bt} collects the results
of the set of models from the two forms of nonlinearity.
Model 1 is the approximation to the Box-Tidwell transformation
from Model 2 of Table \ref{tab:reg_bt_lin}. 
Model 2
has the same specification as the approximate transformation, 
except that the age variable is transformed using the optimal
exponent for the Box-Tidwell transformation. 
% 
The last model has the highest R-squared
among the ones we have estimated, 
with only a slight improvement over the linear approximation.
Again, the differences are marginal, so the practical recommendation
is model with the $f(x) = x\cdot\log(x)$.


\input{../Tables/reg_sq_age_bt}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
